# Rate Books and Rate Calculation Examples {#ratebooks}

In this section we provide rating examples we can reference.

## Homeowners

Homeowners insurance covers damage to the property, contents, and outstanding structures, as well as loss of use, liability and medical coverage. The perils covered and amount of insurance provided is detailed in the policy contract.  

This rate book is based on all-perils combined and all coverages combined for simplicity.

### Homeowners - Rating Algorithm

The rating algorithm is defined as follows.

```{r rating_algorithm, echo = FALSE, warning = FALSE}
tbl_algorithm <-
  tribble(
    ~variable, ~operation, ~rate_table,
    "All-Peril Base Rate", "+", "tbl_base",
    "AOI Relativity", "x", "tbl_aoi",
    "Territory Relativity", "x", "tbl_territory",
    "Protection Class / Construction Type Relativity", "x", "tbl_protection_class",
    "Underwriting Tier Relativity", "x", "tbl_uw_tier",
    "Deductible Credit", "x", "tbl_deductible",
    "New Home Discount", "x [1 - .]", "tbl_credits",
    "Claims-Free Discount", "x [1 - .]", "tbl_credits",
    "Multi-Policy Discount", "x [1 - .]", "tbl_credits",
    "Increased Jewelry Coverage Rate", "+", "tbl_jewelry",
    "Increased Liability/Medical Coverage Rate", "+", "tbl_increased_liability",
    "Policy Fee", "+", "tbl_expense_fee"
  ) 

tbl_algorithm %>% 
  transmute(
    expression = variable %>% paste0(" (", ., ")"),
    expression = 
      if_else(
        operation %>% str_detect("\\."), 
        operation %>% str_replace("\\.", expression), 
        map2_chr(operation, expression, function(x,y) paste0(x, y, collapse = " "))
  ))
```

The rate tables associated with each rating step can be mapped as follows.

```{r rating_algorithm_tabular, echo = FALSE}
tbl_algorithm %>% 
  select(-operation) %>% 
  knitr::kable(format = "html") %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

### Homeowners - Rate Tables

The rate tables are defined as follows.  Note the amount of insurance (`aoi`) table does not typically have a max.  Programming each additional factors (extrapolation) and interpolation can be cumbersome.  At this time, we use rounded `aoi` and use a direct lookup rather than interpolation which is often used in practice.

```{r ho_rate_tables, echo = FALSE}
ho_rate_tables <- 
  list(
    # Base rate
    tbl_base =
      tribble(
        ~peril,     ~value,
        "all_perils", 500),
    tbl_aoi =
      tribble(
        ~aoi,     ~value,
         80 ,0.56
        ,95 ,0.63
        ,110,0.69
        ,125,0.75
        ,140,0.81
        ,155,0.86
        ,170,0.91
        ,185,0.96
        ,200,1.00
        ,215,1.04
        ,230,1.08
        ,245,1.12
        ,260,1.16
        ,275,1.20
        ,290,1.24
        ,305,1.28
        ,320,1.32
        ,335,1.36
        ,350,1.39
        ,365,1.42
        ,380,1.45
        ,395,1.48
        ,410,1.51
        ,425,1.54
        ,440,1.57
        ,455,1.60
        ,470,1.63
        ,485,1.66
        ,500,1.69
        # ,15 ,0.03   Can't do each additional right now.
        ),
    tbl_territory =
      tribble(
        ~territory,     ~value,
        1,0.80,
        2,0.90,
        3,1.00,
        4,1.10,
        5,1.15),
    tbl_protection_class = 
      tribble(
        ~protection_class, ~construction_type, ~value,
        1, "frame",  1.00, 
        2, "frame",  1.00, 
        3, "frame",  1.00, 
        4, "frame",  1.00, 
        5, "frame",  1.05, 
        6, "frame",  1.10, 
        7, "frame",  1.15, 
        8, "frame",  1.25, 
        9, "frame",  2.10, 
        10,"frame",  2.30, 
        1, "masonry",0.90,
        2, "masonry",0.90,
        3, "masonry",0.90,
        4, "masonry",0.90,
        5, "masonry",1.00,
        6, "masonry",1.05,
        7, "masonry",1.10,
        8, "masonry",1.15,
        9, "masonry",1.75,
        10,"masonry",1.90),
    tbl_uw_tier =
      tibble(
        uw_tier = LETTERS[1:10],
        value    = 1:10 * 0.12 + 0.7),
    tbl_deductible =
      tribble(
        ~deductible, ~value,
        250,  1.00,
        500,  0.95,
        1000, 0.85,
        5000, 0.70),
    tbl_credits =
      tribble(
        ~credit,       ~value,
        "new home",    0.2,
        "claims free", 0.1,
        "multipolicy", 0.08),
    tbl_jewelry =
      tribble(
        ~jewelry_limit, ~value,
        2500, 0,
        5000, 35,
        10000, 60),
    tbl_increased_liability =
      tribble(
        ~limit_liability_medpay, ~value,
        "$100,000/$500",   0,
        "$300,000/$1,000", 25,
        "$500,000/$2,500", 45),
    tbl_expense_fee =
      tribble(
        ~form, ~value,
        "HO3", 50)
  )

# ho_rate_tables %>% write_rds("ho_rate_tables.Rds")
```

`r ho_rate_tables %>% map(function(x) knitr::kable(x) %>% kableExtra::kable_styling(c("striped", "condensed"), full_width = FALSE, position = "left"))`

## Rating Calculation Examples

In this section we provide a rating example for each rate book presented as a full-term quoted premium.  

### Homeowners

The policy has the following risk characteristics:

- Amount of insurance = <b>\$275,000</b>.
- The insured lives in Territory <b>2</b>.
- The home is <b>frame</b> construction located in Fire Protection Class <b>8</b>.
- Based on the insured’s credit score, tenure with the company, and prior loss history, the policy has been placed in Underwriting Tier <b>C</b>.
- The insured opts for a <b>\$1,000</b> deductible.
- The home falls under the definition of a <b>new home</b> as defined by the rating rules.
- The insured is eligible for the five-year <b>claims-free</b> discount.
- There is no corresponding auto or excess liability policy.
- The policyholder opts to increase coverage for jewelry to <b>\$5,000</b> and to increase liability/medical coverage limits to <b>\$300,000/$1,000</b>. 

The calculation is expressed as follows.

$500 \times 1.20 \times 0.9 \times 1.25 \times 1.06 \times 0.85 \times (1 - 0.2) \times (1 - 0.1) \times (1 - 0) + 35 + 50 = 522.9$

The final result is rounded to the nearest dollar so the final premium for this quote is <b>\$523</b>.

## Rate Book Complexity

One way to measure the complexity of a rate book is by how many price points are possible.  In practice, this doesn't always mean every scenario will occur or be possible.  For example, age of insured is a widely used rating variable with high predictive power; the rate table may contain values for 120 years olds, but in reality there are no 120 year olds who need insurance.  That being said, the number of possible price points may also measure flexibility of the rating system.  Complexity in this case is referring only to the number of possibilities, not a measure of sophistication.

The number of possible price points for each rate book is as follows:

```{r}
tribble(
  ~rate_book, ~complexity, 
  "Homeowners", ho_rate_tables %>% sapply(nrow) %>% prod
  ) %>% 
kableExtra::kable(format = "html") %>% 
kableExtra::kable_styling(c("striped", "condensed"), full_width = FALSE)
```


