# Generating Premium Data {#premium_data}

To get get started we need data that contains rating characteristics. These rating characteristics (also called features) will be passed through a rate book to generate premiums.  In this section, we create datasets using the rating examples we saw in the last section.  Effectively, we can generate a book of business with any particular mix we like. This will come in handy later when test assumptions and prove the merits of this approach.

## Uniformly Distributed, Independent Mix, with Exact Features

Let's define what is meant by uniformly distributed, independent mix, with exact features.

  - <i>Uniformly Distributed</i>: each level has the same chance of being drawn as any other level.
  - <i>Independent Mix</i>: rating variables are not correlated with one another whatsoever.
  - <i>Exact Features</i>: features are drawn from the rate tables so the exact value can be found in the rate table.

This is the simplest dataset to create.  We randomly sample a row from each rate table and bind the result all together to create the exposure record. This code generates feature data we can use to rate with.  

```{r ho_uniform_iid}
# Number of records we want to generate.
n_draw <- 10000
ho_rate_tables <- read_rds("ho_rate_tables.Rds")

#Set seed to get consistent results each draw.
set.seed(5555) 

# For each table, draw n_draw random numbers.
ho_uniform_iid <-
  ho_rate_tables %>% 
  lapply(
    function(x) {
      nrow(x) %>% 
      sample(n_draw, replace = TRUE) %>% 
      lapply(function(y) x[y, -ncol(x)]) %>% bind_rows}) %>% 
  bind_cols %>% 
  cbind(policy_number = 1:nrow(.) %>% str_pad(nchar(n_draw), pad = "0"), .)

ho_uniform_iid %>% as_tibble
```

Next, retrieve the rating factors and apply them according to the rating algorithm steps to generate premium as follows.  One thing to mention here.  In reality a policy can have multiple credits but this code only randomly draws one from the credit table the way it is setup.  This doesn't take away from our experiment.

```{r}
# Need to use the formula we already laid out above somehow.  This is just to keep going.
ho_uniform_iid_rating <-
  ho_rate_tables %>% 
    lapply(function(x) (x %>% right_join(ho_uniform_iid))$value) %>% 
  bind_rows %>% 
  mutate(
    premium =
      tbl_base *
      tbl_aoi *
      tbl_territory * 
      tbl_protection_class *
      tbl_uw_tier *
      tbl_deductible *
      (1 - tbl_credits) +
      tbl_jewelry +
      tbl_increased_liability +
      tbl_expense_fee,
    premium = premium %>% round
  )

# Inpute premium data onto rating data
ho_uniform_iid <- 
  ho_uniform_iid %>% 
  cbind(premium = ho_uniform_iid_rating$premium) %>% 
  mutate_if(is.character, factor)
```

Let's take a look at the distribution of premiums.

```{r fig.align = 'center'}
ho_uniform_iid %>% 
  ggplot + 
  geom_histogram(aes(x = premium), binwidth = 50, color = "darkorange3", fill = "white") +
  xlab("Premium Bin") +
  ylab("Number of Training Examples") +
  theme_bw()
```

This is great result. An insurance rating plan uses ratings factors typically generated using loss cost models.  Even though we randomly generated a feature set there is a clear transfomation that occurs just by using rating factors selected from generlized linear model.

## Not Uniformly Distributed, Independent Mix of Business

## Not Uniformly Distributed, Dependent Mix of Business

## Generating Historical Book of Business

To this point we have generated datasets that mirror what we'll call a quoting environment rather than a policy writing environment. A quoting environment is only generating full-term premiums based on the characteristics provided.  The policy writing system includes endorsements which can generate different premiums over time if the endorsement is "premium bearing".  The main reason for the differences is simple: time. Quoting is at one point in time with rating features fixed, but a policy can change during it's term.  In other words, the rating features can change over the life of the policy contract for various reasons--- for example, if the insured changes their policy deductible mid-term.

In this section we develop a mock historical book of business.  To keep it simple, we are not going to use new business and retention assumptions.  In other words, we will start with one set of simulated policy data and age it to simulate five years worth of premium data with rate changes introduced.  Policies falling off the book and new business coming on will shift the mix but the \@ref[proof of concept](poc) does not require such detail at this time.  In addition, again for simplicity, we are going to build in a rate change for each each year--- so if we generate five years of data we will include four rate changes.

It is important to note

### Uniformly Distributed

```{r}
ho_uniform_iid$policy_year <- 
  Sys.Date() %>% 
  lubridate::year() - 5

bob_ho_uniform_iid <-
  ho_uniform_iid %>% 
  mutate(
    policy_year = policy_year + 1,
    aoi         = (1.05 * aoi) %>% round
    )
  # lubridate::ceiling_date(unit = "years") - 1
  # lubridate::floor_date(unit = "years")
  # lubridate::year()

# An example of sampling dates
# sample(seq(as.Date('1999/01/01'), as.Date('2000/01/01'), by="day"), nrow(ho_uniform_iid), replace = TRUE)

ho_uniform_iid %>% write_rds("ho_uniform_iid.Rds")
```

### Rate Book Changes



## Extreme Sample Bias

What if 90% of the data was male rather than female.
